---
title: "**Mapping the time-course of in out of consciousness during binocular rivalry: The case of expressions and gender of face stimuli**"
subtitle: "Supplementary Materials"
author: 
  - "Thomas Quettier"
  - "Nicolò Di Lello"
  - "Naotsugu Tsuchiya"
  - "Paola Sessa"
output:
  bookdown::pdf_document2:
    toc: true
    toc_depth: 3
    highlight: pygments
bibliography: ["../files/references.bib"]
header-includes:
  - \AddToHook{cmd/section/before}{\clearpage}
  - \usepackage{titling}
  - \setcounter{table}{0} 
  - \renewcommand*{\thetable}{S\arabic{table}}
  - \setcounter{figure}{0} 
  - \renewcommand*{\thefigure}{S\arabic{figure}}
  - \usepackage{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \usepackage[nottoc]{tocbibind}
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center",
                      fig.retina = 2,
                      dev = "pdf")
```

```{r packages, cache=FALSE}

library(here)
library(magrittr)
library(tidyr)
library(dplyr)
library(kableExtra)
library(flextable)
library(ggplot2)
library(emmeans)
library(lme4)
library(lmerTest)
library(effects)
library(tidyverse)
library(xtable)
library(ftExtra)
library(broom)
library(broom.mixed)
library(effects)
library(ggcorrplot)
library(PerformanceAnalytics)

devtools::load_all()
```

```{r data}
dat <- readRDS(here("data","data.rds"))
emo <- readRDS(here("data","emotion_correlation.rds"))
gen <- readRDS(here("data","gender_correlation.rds"))
stab <- readRDS(here("data","stability.rds"))
```


# General approach

We used R (R Core Team, 2012) and lme4 (Bates, Maechler & Bolker, 2012) to perform a linear mixed effects analysis of the relationship between rivalry, consciousness phases  and subjects gender. As fixed effects, we entered rivalry and phases and subject gender (with interaction term) into the model. As random effects, we had intercepts for subjects. Visual inspection of residual plots did not reveal any obvious deviations from homoscedasticity or normality. P-values were obtained by anova of the full models for gender blocks and emotion blocks:


```{r bestmodel, eval = FALSE, echo = TRUE}

fit <- lmer(mean ~ rivalry*phase*gender + (1|subject)

```

\newpage
## model selection

The logic of the model selection is to compare the likelihood of different models. First, the model without any factor (the
null model), then each model add a factor that we are interested in.

```{r modelsselection, eval = FALSE, echo = TRUE}

model1<-lmer(speed ~ 1 + (1|subject), dat) # null model

model2<-lmer(speed ~ phase + (1|subject), data) # add consciousness phases: formation vs dissolution

model3<-lmer(speed ~ phase + rivalry + (1|subject), data) # add rivalry: emotion(happy vs neutral) OR gender(male vs female)

model4<-lmer(speed ~ rivalry*phase + (1|subject), data) # add interaction


anova(model1,model2,model3,model4)

# Stabilisation

model1<-lmer(STB ~ 1 + (1|subject), dat) # null model

model2<-lmer(STB ~ rivalry + (1|subject), data) # add consciousness rivalry: emotion(happy vs neutral) OR gender(male vs female)

anova(model1,model2)

```


### Model selection : Emotion rivalry

```{r anovamodels_emotion, results='asis'}
data <- dat%>%filter(procedure=="emotion")

model1<-lmer(mean ~ 1 + (1|subject), data)
model2<-lmer(mean ~ phase + (1|subject), data)
model3<-lmer(mean ~ phase + rivalry + (1|subject), data)
model4<-lmer(mean ~ rivalry*phase + (1|subject), data , contrasts = list(phase = contr.sum(2)/2, rivalry = contr.sum(2)/2))

modelselection<- anova(model1,model2,model3,model4)

modelselection%>%
  tidy()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
  width( width = 0.8)%>%
   # set_table_properties(layout = "autofit")%>%
      theme_vanilla()%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")
```



### Model selection : Gender rivalry


```{r anovamodels_gender, results='asis'}
data <- dat%>%filter(procedure=="gender")

model1<-lmer(mean ~ 1 + (1|subject), data)
model2<-lmer(mean ~ phase + (1|subject), data)
model3<-lmer(mean ~ phase + rivalry + (1|subject), data)
model4<-lmer(mean ~ rivalry*phase + (1|subject), data , contrasts = list(phase = contr.sum(2)/2, rivalry = contr.sum(2)/2))

modelselection<- anova(model1,model2,model3,model4)
modelselection%>%
  tidy()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
  width( width = 0.8)%>%
      theme_vanilla()%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


```

\newpage
### Model selection : Emotion stability

```{r anovamodels_stabemo, results='asis'}
data <- stab%>%filter(procedure=="emotion")
model1<-lmer(CT ~ 1 + (1|subject), data)
model2<-lmer(CT ~ rivalry + (1|subject), data)



modelselection<- anova(model1,model2)

modelselection%>%
  tidy()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
  width( width = 0.8)%>%
      theme_vanilla()%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


```
### Model selection : gender stability

```{r anovamodels_stabgen, results='asis'}
data <- stab%>%filter(procedure=="gender")
model1<-lmer(CT ~ 1 + (1|subject), data)
model2<-lmer(CT ~ rivalry + (1|subject), data)



modelselection<- anova(model1,model2)

modelselection%>%
  tidy()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
  width( width = 0.8)%>%
      theme_vanilla()%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


```



# Emotion Rivalry

## Anova model 4 emotion

```{r anova_emotion, results='asis'}
data <- dat%>%filter(procedure=="emotion")
model4<-lmer(mean ~ rivalry*phase + (1|subject), data)

fit<- anova(model4)


fit%>%
  tidy()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


# plot(allEffects(fit))

```
```{r plot_emotion}

 plot(allEffects(model4))

```
Warning: Speed mean that higher values are faster.

\newpage

### Contrasts main effect phases
```{r contrast_emotion_phase, results='asis'}
x<-emmeans(model4,pairwise~ phase, adjust="bonf")

x$emmeans%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Means phases emotion")


x$contrasts%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Contrasts phases emotion")%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


```
Note: phases are asymmetric: Dissolution is faster than formation.



### Contrasts interaction rivalry:phases
```{r contrast_emotion_rivalryphase, results='asis'}
x<-emmeans(model4,pairwise~ rivalry|phase, adjust="bonf")

x$emmeans%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Means rivalry|phases emotion")


x$contrasts%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Contrasts rivalry|phases emotion")%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


```
Notes: Happy Dissolution is faster than neutral Dissolution.
Neutral Formation is faster than happy Formation.


\newpage
# Gender Rivalry

## Anova model 2 gender

```{r anova_gender, results='asis'}
data <- dat%>%filter(procedure=="gender")
model2<-lmer(mean ~ phase + (1|subject), data)

fit<- anova(model2)


fit%>%
  tidy()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


# plot(allEffects(fit))

```
```{r plot_gender}

 plot(allEffects(model2))

```
Warning: Speed mean that higher values are faster.

\newpage

### Contrasts main effect phases
```{r contrast_gender_phase, results='asis'}
x<-emmeans(model2,pairwise~ phase, adjust="bonf")

x$emmeans%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Means phases gender")


x$contrasts%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Contrasts phases gender")%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


```
Note: phases are asymmetric: Dissolution is faster than formation.




\newpage
# Stability

## Anova stability emotion model 2 
```{r anova_emotion_stability, results='asis'}
data <- stab%>%filter(procedure=="emotion")
model2<-lmer(CT ~ rivalry + (1|subject), data )

fit<- anova(model2)


fit%>%
  tidy()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


# plot(allEffects(fit))

```
```{r plot_emotion_stability}

 plot(allEffects(model2))

```
Warning: CT mean that higher values are more stable.


### Contrasts main effect rivalry
```{r contrast_emotion_rivalry_stability, results='asis'}
x<-emmeans(fit,pairwise~ rivalry, adjust="bonf")

x$emmeans%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Means rivalry emotion")


x$contrasts%>%
  as.data.frame()%>%
  flextable()%>%
  colformat_double(digits = 3) %>% 
    autofit()%>%
      theme_vanilla()%>%
  add_header_lines(values = "Contrasts rivalry emotion")%>%
  bg( i = ~ p.value < 0.05, bg = "#e6cc77", part = "body")


```





